\Large{Research Seminar Software Science} \\ \huge{Simulation of monadic effects in Haskell}

### 2018-06-02 paper.tex
created bridge between problem and solution for plus parser

    \begin{tcolorbox}
    \begin{lstlisting}[language=Haskell]
      (+++) :: Parse a -> Parse a -> Parse a
      p +++ q  = \input -> (p input ++ q input)
    \end{lstlisting}
    \end{tcolorbox}
    \noindent The nature of the plus parser can be illustrated by defining the \textbf{islower\_r} parser which alternated two of the already defined parsers from the previous subsection: \textbf{islowercase} and \textbf{singleChar}.

@@

    \noindent Now, let's apply the \textbf{islower\_r} parser to an input string that contains `r' as it's first character.

@@

     With this modification, some deterministic behaviour has be inserted in the \textbf{plus} parser.

>>>>

    \begin{tcolorbox}
    \begin{lstlisting}[language=Haskell]
      (+++) :: Parse a -> Parse a -> Parse a
      p +++ q  = \input -> (p input ++ q input)
    \end{lstlisting}
    \end{tcolorbox}
    \noindent The nature of the plus parser can be illustrated by defining the \textbf{isletter} parser which alternates two of the already defined parsers from the previous subsection: \textbf{islowercase} and \textbf{isuppercase}.

    \begin{tcolorbox}
    \begin{lstlisting}[language=Haskell] 
        isletter :: Parse Char
        isletter  = islowercase +++ isuppercase
    \end{lstlisting}
    \end{tcolorbox}

    \noindent The \textbf{isletter} parser will parse the first character of input string if it is a lowercase or an uppercase letter. How does it accomplish this? Firstly, it will apply parser \textbf{islowercase} to the input string and yield a tuple with resulting parsed character together with remaining not parsed input string. Secondly, it will apply the parser \textbf{isuppercase} to the same input string and return a tuple containing the resulting  parsed character together with the yet to be parsed input string. Finally, it will concatenate the resulting tuples from \textbf{islowercase} and \textbf{isuppercase} parsers into one list. As shown from the example below, the \textbf{isletter} parser is exhibiting the correct behavior by just returning the parsed character is it is a letter.

    \begin{center}
            \textbf{ghci$>$} isletter \textcolor{red}{``rera"} \\
            $\big[$(\textcolor{green}{`r'}, \textcolor{red}{``era"})$\big]$
    \end{center}

    \noindent Actually, the resulting outcome is the output produced by the \textbf{islowercase} parser while the \textbf{isuppercase} parser returns an empty list. If the \textbf{isuppercase} parser returned a non empty list , the resulting output would be appended to the output of the \textbf{islowercase} parser, thus by producing some ambiguous behavior. In other words, there is an issue with the non-deterministic behavior of the \textbf{plus} parser. \\ \\
    The nature of this complication can be clearly represented by the \textbf{islower\_r} parser which alternates two of the already defined parsers from the previous subsection: \textbf{islowercase} and \textbf{singleChar}.

@@

    \noindent The role that the \textbf{islower\_r} exhibits is not very useful in parsing terms but in this case it is essential in exposing the hidden inadequate behavior of the \textbf{plus} parser. Now, let's apply the \textbf{islower\_r} parser to an input string that contains `r' as it's first character to observe what happens.

@@

     With this modification, some deterministic behaviour has been inserted in the \textbf{plus} parser.
